---
title: "Trends"
output:
  html_document:
    theme: flatly
    self_contained: true
params:
  data_path: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(plotly)
library(jsonlite)

d <- readRDS(params$data_path)

# Everything the page needs travels as one small JSON array (one object per
# activity). All windowing, bucketing and per-sport aggregation happens
# client-side in the script below, so the controls re-aggregate instantly with
# no server and no re-render. Keeping it here (rather than pre-computing every
# control combination in R) is what makes an arbitrary multi-select of sports
# cheap: the JS sums the raw rows on demand.
payload <- jsonlite::toJSON(
  d, dataframe = "rows", na = "null", auto_unbox = TRUE
)
```

<style>
#trends-controls { margin: 0 0 1rem 0; display: flex; flex-wrap: wrap;
  gap: 1.25rem; align-items: flex-end; }
#trends-controls .grp { display: flex; flex-direction: column; gap: 0.2rem; }
#trends-controls label.hdr { font-size: 0.75rem; text-transform: uppercase;
  letter-spacing: 0.03em; color: #666; }
#trends-controls select { padding: 0.25rem 0.4rem; }
#trends-sports { display: flex; flex-wrap: wrap; gap: 0.1rem 0.9rem;
  max-width: 640px; }
#trends-sports label.trends-chk { font-weight: normal; margin: 0;
  white-space: nowrap; }
#trends-sport-btns { display: flex; gap: 0.4rem; }
#trends-sport-btns button { font-size: 0.75rem; padding: 0.1rem 0.5rem; }
#trends-chart { width: 100%; height: 520px; }
</style>

<div id="trends-controls">
  <div class="grp">
    <label class="hdr" for="trends-metric">Metric</label>
    <select id="trends-metric">
      <option value="distance" selected>Distance (km)</option>
      <option value="count">Activity count</option>
      <option value="elapsed">Elapsed time (h)</option>
      <option value="moving">Moving time (h)</option>
    </select>
  </div>
  <div class="grp">
    <label class="hdr" for="trends-bucket">Bucket</label>
    <select id="trends-bucket">
      <option value="week">Week</option>
      <option value="month" selected>Month</option>
    </select>
  </div>
  <div class="grp">
    <label class="hdr" for="trends-period">Period (from newest)</label>
    <select id="trends-period">
      <option value="1m">1 month</option>
      <option value="1y" selected>1 year</option>
      <option value="3y">3 years</option>
      <option value="5y">5 years</option>
      <option value="10y">10 years</option>
      <option value="all">All time</option>
    </select>
  </div>
  <div class="grp">
    <label class="hdr">Sports</label>
    <div id="trends-sport-btns">
      <button type="button" id="trends-all">All</button>
      <button type="button" id="trends-none">None</button>
    </div>
    <div id="trends-sports"></div>
  </div>
</div>

```{r chart, echo=FALSE}
# An empty widget solely to bundle plotly.js into this self-contained page and
# provide a graph div with a known id. The script below fills it via
# Plotly.react; the empty plot is never shown.
p <- plotly::plot_ly()
p$elementId <- "trends-chart"
p
```

```{r payload, results="asis"}
cat(sprintf("<script>window.STARCH_TRENDS = %s;</script>", payload))
```

<script>
(function () {
  var DATA = window.STARCH_TRENDS || [];
  var METRIC_LABEL = {
    count: "Activities",
    distance: "Distance (km)",
    elapsed: "Elapsed time (h)",
    moving: "Moving time (h)"
  };

  // ---- aggregation core (unit-tested separately) --------------------------

  function parseUTC(s) {
    var p = s.split("-");
    return new Date(Date.UTC(+p[0], +p[1] - 1, +p[2]));
  }
  function pad2(n) { return (n < 10 ? "0" : "") + n; }
  function ymd(dt) {
    return dt.getUTCFullYear() + "-" + pad2(dt.getUTCMonth() + 1) + "-" +
      pad2(dt.getUTCDate());
  }
  // Monday of the activity's week (Monday-start), in UTC.
  function weekStart(dt) {
    var d = new Date(dt.getTime());
    var day = d.getUTCDay();          // 0=Sun..6=Sat
    var iso = (day === 0) ? 7 : day;  // 1=Mon..7=Sun
    d.setUTCDate(d.getUTCDate() - (iso - 1));
    return d;
  }
  function monthStart(dt) {
    return new Date(Date.UTC(dt.getUTCFullYear(), dt.getUTCMonth(), 1));
  }
  function bucketStart(dt, bucket) {
    return bucket === "week" ? weekStart(dt) : monthStart(dt);
  }
  function bucketNext(dt, bucket) {
    var d = new Date(dt.getTime());
    if (bucket === "week") { d.setUTCDate(d.getUTCDate() + 7); }
    else { d.setUTCMonth(d.getUTCMonth() + 1); }
    return d;
  }
  // Window lower bound: anchor (newest activity) shifted back.
  function windowStart(anchor, period, minDate) {
    var y = anchor.getUTCFullYear(), m = anchor.getUTCMonth(),
      dd = anchor.getUTCDate();
    switch (period) {
      case "1m": return new Date(Date.UTC(y, m - 1, dd));
      case "1y": return new Date(Date.UTC(y - 1, m, dd));
      case "3y": return new Date(Date.UTC(y - 3, m, dd));
      case "5y": return new Date(Date.UTC(y - 5, m, dd));
      case "10y": return new Date(Date.UTC(y - 10, m, dd));
      default: return minDate; // "all"
    }
  }
  function metricValue(row, metric) {
    if (metric === "count") return 1;
    var v = metric === "distance" ? row.distance_km
      : metric === "elapsed" ? row.elapsed_s
        : row.moving_s;
    if (v === null || v === undefined || isNaN(v)) return 0;
    return (metric === "distance") ? v : v / 3600; // seconds -> hours
  }
  // Sports present, ordered by descending all-history activity count.
  function sportOrder(rows) {
    var counts = {};
    rows.forEach(function (r) { counts[r.type] = (counts[r.type] || 0) + 1; });
    return Object.keys(counts).sort(function (a, b) {
      return counts[b] - counts[a] || (a < b ? -1 : a > b ? 1 : 0);
    });
  }
  function aggregate(rows, opts) {
    var bucket = opts.bucket, metric = opts.metric,
      sports = opts.sports, period = opts.period;

    var parsed = rows.map(function (r) {
      return { d: parseUTC(r.date), type: r.type, row: r };
    });
    if (parsed.length === 0) return { buckets: [], series: {} };

    var anchor = parsed.reduce(function (mx, p) { return p.d > mx ? p.d : mx; },
      parsed[0].d);
    var minAll = parsed.reduce(function (mn, p) { return p.d < mn ? p.d : mn; },
      parsed[0].d);

    var cutoff = windowStart(anchor, period, minAll);
    var sel = {};
    sports.forEach(function (s) { sel[s] = true; });

    var filtered = parsed.filter(function (p) {
      return p.d >= cutoff && sel[p.type];
    });
    if (filtered.length === 0) return { buckets: [], series: {} };

    var lo = filtered.reduce(function (mn, p) { return p.d < mn ? p.d : mn; },
      filtered[0].d);
    var hi = filtered.reduce(function (mx, p) { return p.d > mx ? p.d : mx; },
      filtered[0].d);

    // Contiguous bucket sequence from first to last in-window activity, so
    // inactive weeks/months appear as gaps rather than being skipped.
    var seq = [];
    var cur = bucketStart(lo, bucket), end = bucketStart(hi, bucket);
    while (cur <= end) { seq.push(ymd(cur)); cur = bucketNext(cur, bucket); }
    var idx = {};
    seq.forEach(function (k, i) { idx[k] = i; });

    var series = {};
    sports.forEach(function (s) {
      if (sel[s]) series[s] = seq.map(function () { return 0; });
    });
    filtered.forEach(function (p) {
      var key = ymd(bucketStart(p.d, bucket));
      series[p.type][idx[key]] += metricValue(p.row, metric);
    });
    return { buckets: seq, series: series };
  }

  // ---- controls + rendering -----------------------------------------------

  function el(id) { return document.getElementById(id); }
  var ORDER = [];   // full sport order, count descending
  var GD = null;    // graph div, resolved once

  function chartDiv() {
    return GD || el("trends-chart") ||
      document.querySelector(".plotly.html-widget");
  }
  function selectedSports() {
    var out = [];
    document.querySelectorAll("#trends-sports input[type=checkbox]")
      .forEach(function (b) { if (b.checked) out.push(b.value); });
    return out;
  }
  function buildSportControls(order) {
    var host = el("trends-sports");
    host.innerHTML = "";
    order.forEach(function (s) {
      var lab = document.createElement("label");
      lab.className = "trends-chk";
      var box = document.createElement("input");
      box.type = "checkbox"; box.value = s; box.checked = true;
      box.addEventListener("change", render);
      lab.appendChild(box);
      lab.appendChild(document.createTextNode(" " + s));
      host.appendChild(lab);
    });
  }
  function setAll(v) {
    document.querySelectorAll("#trends-sports input[type=checkbox]")
      .forEach(function (b) { b.checked = v; });
    render();
  }
  function render() {
    var metric = el("trends-metric").value;
    var agg = aggregate(DATA, {
      bucket: el("trends-bucket").value,
      period: el("trends-period").value,
      metric: metric,
      sports: selectedSports()
    });
    // Traces follow the fixed count-descending order, so a sport keeps its
    // colour and stacking position as the selection changes.
    var traces = ORDER.filter(function (s) { return agg.series[s]; })
      .map(function (s) {
        return { type: "bar", name: s, x: agg.buckets, y: agg.series[s] };
      });
    var layout = {
      barmode: "stack",
      height: 520,
      autosize: true,
      margin: { l: 60, r: 20, t: 20, b: 40 },
      bargap: 0.05,
      xaxis: { type: "date" },
      yaxis: { title: { text: METRIC_LABEL[metric] }, rangemode: "tozero" },
      legend: { orientation: "h" }
    };
    var gd = chartDiv();
    if (gd && typeof Plotly !== "undefined") {
      GD = gd;
      Plotly.react(gd, traces, layout, { responsive: true, displaylogo: false });
    }
  }
  function init() {
    ORDER = sportOrder(DATA);
    buildSportControls(ORDER);
    ["trends-metric", "trends-bucket", "trends-period"].forEach(function (id) {
      el(id).addEventListener("change", render);
    });
    el("trends-all").addEventListener("click", function () { setAll(true); });
    el("trends-none").addEventListener("click", function () { setAll(false); });
    render();
  }
  // Defer past htmlwidgets' own on-load init (which draws the empty plot) and
  // wait for plotly.js to be present, so our first react is the one that wins.
  function boot() {
    if (typeof Plotly === "undefined") { return setTimeout(boot, 30); }
    init();
  }
  if (document.readyState === "complete") { setTimeout(boot, 0); }
  else { window.addEventListener("load", function () { setTimeout(boot, 0); }); }
})();
</script>
